---
title: "R Notebook"
output: html_notebook
---

# Load Required Libraries

```{r}
library(dplyr)
library(MASS)
library(naniar) 
library(ROCR) 
library(rpart)
library(knitr)
library(tibble)
library(caret)
library(gridExtra)
library(car)
library(tseries)
library(lmtest)
library(nortest)
```
# Data Analysis

## Load and Shuffle the Data

```{r}
set.seed(12345) 
data <- read.csv("WineQT.csv")
data <- dplyr::select(data, -"Id")  
data <- data[sample(nrow(data)), ]  
```

## Summarize the Dataset

```{r}
summary(data)
```

```{r}
GGally::ggcorr(
  data, 
  method = c("complete", "pearson"), 
  label = TRUE, 
  geom = "tile", 
  label_round = 3, 
  label_size = 2, 
  legend.size = 6, 
  size = 1.5, 
  hjust = 0.75
)
```

# Basic Model Building

## Initial Model

```{r}
model1<-lm(quality~.,data=data)
summary(model1)
```

## Iterative Feature Selection
The model is refined step by step by removing predictors one at a time.

```{r}
model1<-lm(quality~.-density,data=data)
summary(model1)
```

```{r}
model1<-lm(quality~.-density-fixed.acidity,data=data)
summary(model1)
```
```{r}
model1<-lm(quality~.-density-fixed.acidity-residual.sugar,data=data)
summary(model1)
```
```{r}
model1<-lm(quality~.-density-fixed.acidity-residual.sugar-citric.acid,data=data)
summary(model1)
```
```{r}
model1<-lm(quality~.-density-fixed.acidity-residual.sugar-citric.acid-free.sulfur.dioxide,data=data)
summary(model1)
```


```{r}
model1<-lm(quality~volatile.acidity+total.sulfur.dioxide+pH+sulphates+alcohol+chlorides:sulphates+chlorides,data=data)
summary(model1)
```
## Normality of Residuals

### Normality Tests

```{r}
jarque.bera.test(model1$residuals)
```
```{r}
lillie.test(model1$residuals)
```
```{r}
ad.test(model1$residuals)
```

We can see issues with residuals normality, however when analysing the histogram: 

```{r}
hist(model1$residuals, freq=FALSE,breaks=20)
lines(seq(from = min(model1$residuals), to = max(model1$residuals),by=0.01),dnorm(seq(from = min(model1$residuals), to = max(model1$residuals),by=0.01),sd=sd(model1$residuals)))
```
```{r}
mean(model1$residuals)
```

We can see it resembles normal distribution. Moreover, let's point out we are dealing with some outliers in the training set.

```{r}
plot(model1,which=4)
```

We will get rid of them to validate, whether our model isn't affected by them.

# Model without outliers

```{r}
remove_high_leverage <- function(model, threshold = 4/(nrow(model$model) - length(model$coefficients))) {

  cooks_distance <- cooks.distance(model)

  high_leverage_indices <- which(cooks_distance > threshold)

  cleaned_data <- model$model[-high_leverage_indices, , drop = FALSE]
  
  cleaned_model <- update(model, data = cleaned_data)
  
  return(list(model = cleaned_model,data=cleaned_data))
}

model1_nolev <- remove_high_leverage(model1)$model
cleaned_data <- remove_high_leverage(model1)$data
```

```{r}
summary(model1_nolev)
```
## Normality of Residuals After Removal
```{r}
jarque.bera.test(model1_nolev$residuals)
```

As we can see after removing outliers we can confirm with 5% level of significance that residuals of our model have normal distribution.


## Homoskedasticity Tests

```{r}
ncvTest(model1_nolev)
```
```{r}
gqtest(model1_nolev)
```
```{r}
hmctest(model1_nolev)
```
## Autocorrelation Tests
```{r}
bgtest(model1_nolev)
```
```{r}
dwtest(model1_nolev)
```
## Graphical Analysis

```{r}
plot(model1_nolev)
```
```{r}
qqPlot(model1_nolev)
```

# Box-Cox Transformation

```{r}
lambda <- MASS::boxcox(model1_nolev)$x[which.max(MASS::boxcox(model1_nolev)$y)]
```


```{r}
cleaned_data$trans_quality <- log(cleaned_data$quality)

model2 <- lm(trans_quality~volatile.acidity+total.sulfur.dioxide+pH+sulphates+alcohol+chlorides:sulphates+chlorides,data=cleaned_data)
```

```{r}
summary(model2)
```

```{r}
hist(model2$residuals,freq=FALSE)
lines(seq(from=min(model2$residuals),to=max(model2$residuals),by=0.01),dnorm(seq(from=min(model2$residuals),to=max(model2$residuals),by=0.01),sd=sd(model2$residuals)))
```

## Normality Tests for Transformed Model
```{r}
jarque.bera.test(model2$residuals)
```
```{r}
qqPlot(model2)
```


## Homoskedasticity Tests

```{r}
ncvTest(model2)
```

```{r}
gqtest(model2)
```
```{r}
hmctest(model2)
```

## Autocorrelation Tests
```{r}
bgtest(model2)
```
```{r}
dwtest(model2)
```



